---
name: Dylan Manfre and Natalie Davis
title: "R Notebook"
output: html_notebook
---

Notes
- Link to the data I downloaded from PG Open Data https://data.princegeorgescountymd.gov/Public-Safety/Crime-Incidents-February-2017-to-5th-July-2023/wb4e-w4nf

library
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages('tidygeocoder')
library(tidygeocoder)
#install.packages("curl")
library(curl)
library(ggplot2)
library(ggrepel)
```

loading data
```{r}
pg_county_crime_old <-  read.csv("data/pg_county_crime.csv")
moco_data <- read_csv("data/test_data.csv")
pg_county_crime_new <-  read.csv("data/pg_county_crime_new.csv")

```

______

cleaning data
```{r}
pg_county_crime_old <- pg_county_crime_old %>% 
  clean_names()

pg_county_crime_new <- pg_county_crime_new %>% 
  clean_names()

```
making the dates appropreate
```{r}

#extracting the year column and making the date the correct format
pg_county_crime_old <- pg_county_crime_old %>%
 mutate(date = as.Date(date, "%m/%d/%Y"),
        year = year(date),
         month = month(date)) %>% 
  filter(date <= "2023-07-01")

pg_county_crime_new <- pg_county_crime_new %>% 
  mutate(date = as.Date(date, "%m/%d/%Y"),
        year = year(date),
         month = month(date))


# Renaming the crime column to be easier to work with
pg_county_crime_old <-  pg_county_crime_old %>% 
  rename(crime_type = clearance_code_inc_type)

pg_county_crime_new <- pg_county_crime_new %>% 
   rename(crime_type = clearance_code_inc_type)


```

Showing distinct crimes: 22 types of crime here.
```{r}
pg_county_crime_old %>% 
  distinct(crime_type)

pg_county_crime_new %>% 
  distinct(crime_type)
```
Ditching records before 2020. We are doing this because we felt records 2020 to 2023 provided a good average.
```{r}

pg_county_crime_old <- pg_county_crime_old %>% 
  filter(year >= 2019)

#pg county_crime_new from July 5 - Oct. 10

#binding

total_pg_county_data <- bind_rows(pg_county_crime_old, pg_county_crime_new )

```


----- Exploratory data section -------

What I want to accomplish here?
- I want to see monthly data that goes back a few years for stolen, auto. DONE

filtering for AUTO, STOLEN
```{r}
pg_stolen_auto <- total_pg_county_data %>% 
  filter(crime_type == "AUTO, STOLEN" | crime_type == "AUTO, STOLEN & RECOVERED" )
```
Time elements
```{r}
# yearly data
pg_yearly_data <- pg_stolen_auto %>%
  group_by(year) %>%
  summarize(stolen_auto = n()) %>% 
  mutate(perc = (stolen_auto / sum(stolen_auto)) * 100)

```
Confirms a steady increase from 2020 to 2023. 
Showing on a graph to visualize the increase.
```{r}
ggplot() + 
  geom_bar(
    data=pg_yearly_data, 
    aes(x=year, weight=`stolen_auto`))+
  labs(title = "Car thefts in PG county",
       caption =  "Source: PG Open Data | Graph by Dylan Manfre")

```

monthly totals

```{r}
# What I did here first was make individual month totals by year. I thought there was a condenced way to do this so I used Chat GPT to write this look to iterate over each year from 2017 to 2023. 

# My next attempt was at making a loop before Derek suggested making a month/year column

pg_month_total <- pg_stolen_auto %>% 
  group_by(month, year) %>% 
  summarize(autos_stolen = n()) %>% 
  mutate(month = month(month, label = TRUE))
  #pivot_wider(names_from = year, values_from = count)

write_csv(pg_month_total, "data/pg_monthly_total.csv")

```
auto stolen and recovered notes: There were 782 cars stolen and recovered of the 11,744 which is roughly 7%
```{r}
pg_stolen_recovered_only <- total_pg_county_data %>% 
  filter(crime_type == "AUTO, STOLEN & RECOVERED")

#What percentage were stolen
pg_perc_stolen <- pg_stolen_auto %>% 
  group_by(year, crime_type) %>% 
  summarize(count = n()) %>% 
  #mutate(perc = (count / sum(count)) * 100) %>% 
    pivot_wider(names_from = year, values_from = count)



```
yearly total for stolen and recovered
```{r}

pg_stolen_recovered_years <- pg_stolen_recovered_only %>%
  group_by(year) %>%
  summarize(stolen_recovered = n()) %>%   arrange(desc(stolen_recovered))

pg_stolen_recovered_months <- pg_stolen_recovered_only %>%
  group_by(month, year) %>%
  summarize(stolen_recovered = n()) %>% 
   mutate(month = month(month, label = TRUE)) %>%
  arrange(desc(stolen_recovered))

#joining pg_stolen_recovered_months with pg_month_total

combined_pg_monthly <- inner_join(pg_stolen_recovered_months, pg_month_total, by=c("month", "year"))

write_csv(combined_pg_monthly, "data/combined_pg_monthly.csv")

```

```{r}
#join pg_stolen_recovered_years and #pg_yearly_data

combined_pg_years <- inner_join(pg_stolen_recovered_years, pg_yearly_data, by=c("year")) %>% 
  select(-perc)

```



test <-  stolen_auto %>% 
  filter(incident_case_id == "PP21062600000290")


-----Reverse geocoding section for PG County -----

I need to break up the stolen_auto data into three sections.
Then I need to run the reverse_geocode.
Then I need to bind them all together.


#breaking up stolen auto
#breaking it up into 4 subsets with 2,936 rows because 11,744/4 = 2,936
#Code was copied from ChatGPT

# Number of rows you want in each subset
subset_size <- 1468

# Split the data into eight subsets
subset1 <- stolen_auto[1:subset_size,]
subset2 <- stolen_auto[(subset_size + 1):(2 * subset_size), ]
subset3 <- stolen_auto[(2 * subset_size + 1):(3 * subset_size), ]
subset4 <- stolen_auto[(3 * subset_size + 1):(4 * subset_size), ]
subset5 <- stolen_auto[(4 * subset_size + 1):(5 * subset_size), ]
subset6 <- stolen_auto[(5 * subset_size + 1):(6 * subset_size), ]
subset7 <- stolen_auto[(6 * subset_size + 1):(7 * subset_size), ]
subset8 <- stolen_auto[(7 * subset_size + 1):(8 * subset_size), ]



reverse geocoding

#subset2 <- subset2 %>%
  reverse_geocode(lat = latitude, long = longitude, method = 'osm',
                  address = street_address, full_results = TRUE)


_______

MoCo Data

```{r}
# cleaning

moco_data <- moco_data %>%
   clean_names() %>% 
  separate(start_date_time, c('start_date', 'start_time'), sep=' ', extra='merge') %>%
  mutate(start_date = as.Date(start_date, format = "%m/%d/%Y"),
          year = year(start_date),
         month = month(start_date)) %>% 
  filter(year >= 2019) %>% 
filter(crime_name2 == "Motor Vehicle Theft") %>% 
  rename(crime_name = crime_name2)

moco_data %>% 
  filter(year == 2023) %>% 
  group_by(city) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))

moco_data <- moco_data %>% 
  mutate(city = case_when(
    city == "SILVE SPRING" ~ "SILVER SPRING",
    city == "BEHTESDA" ~ "BETHESDA",
    city == "SILVERS SPRING" ~ "SILVER SPRING",
    city == "GAIHTERSBURG" ~ "GAITHERSBUG",
    city == "BETHEDA" ~ "BETHESDA",
    city == "BARNSVILLE" ~ "BARNESVILLE",
    city == "BEALSVILLE" ~ "BEALLSVILLE",
    city == "CLAEKSBURG" ~ "CLARKSBURG",
    city == "GAIHTERSBURG" ~ "GAITHERSBURG",
    city == "GAITHERBURG" ~ "GAITHERSBURG",
    city == "GAITHERSBUG" ~ "GAITHERSBURG",
    city == "GAUTHERSBURG" ~ "GAITHERSBURG",
    city == "GAITHERSBUG" ~ "GAITHERSBURG",
    city == "GRMANTOWN" ~ "GERMANTOWN",
    city == "KENSTINGTON" ~ "KENSINGTON",
    city == "MT AIRY" ~ "MOUNT AIRY",
    city == "TACOMA PARK" ~ "TAKOMA PARK",
    city == "WEHATON" ~ "WHEATON",
    TRUE ~ city
  ))

```

------------ EXPLORATORY DATA FOR MOCO -----------

Iso 2023
```{r}
moco_2023 <- moco_data %>% 
  filter(year == "2023", crime_name == "Motor Vehicle Theft")

#exporting 2023

write_csv(moco_2023, "data/moco_2023.csv")

```

All moco car thefts
```{r}
moco_car_thefts <- moco_data %>% 
  filter(crime_name == "Motor Vehicle Theft")

write_csv(moco_car_thefts, "data/moco_car_thefts.csv")

View(moco_car_thefts)
```

Quick visual of the yearly thefts
```{r}
moco_yearly <- moco_car_thefts %>% 
  group_by(year) %>% 
  summarize(count = n()) %>% 
  mutate(perc = count/sum(count)*100)

ggplot() + 
  geom_bar(
    data=moco_yearly, 
    aes(x=year, weight=`count`))+
  labs(title = "Montgomery County car thefts",
       caption =  "Source: MoCo Open Data | Graph by Dylan Manfre")

write_csv(moco_yearly, "data/moco_yearly.csv")


```
The majority of the thefts happened in residential parking lots... roughly 29%.
```{r}
moco_theft_places <- moco_car_thefts %>% 
  group_by(place) %>% 
  summarize(count = n()) %>% 
  mutate(perc = count/sum(count)*100) %>% 
  arrange(desc(count))

#2023 only

moco_car_thefts %>% 
  filter(year == 2023) %>% 
  group_by(place) %>% 
  summarize(count = n()) %>% 
  mutate(perc = count/sum(count)*100) %>% 
  arrange(desc(count))

```
Moco monthly totals 
```{r}
 moco_monthly <- moco_car_thefts %>% 
  group_by(month, year) %>% 
  summarize(count = n()) %>% 
  mutate(month = month(month, label = TRUE)) %>% 
  pivot_wider(names_from = year, values_from = count, values_fill = 0)
```





