library
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages('tidygeocoder')
library(tidygeocoder)
#install.packages("curl")
library(curl)
library(ggplot2)
library(ggrepel)
```

loading data
```{r}
pg_county_crime_old <-  read.csv("data/pg_county_crime.csv")
```

cleaning
```{r}
pg_county_crime_old <- pg_county_crime_old %>% 
  clean_names() %>% 
 mutate(date = as.Date(date, "%m/%d/%Y"),
        year = year(date),
         month = month(date)) %>% 
  rename(crime_type = clearance_code_inc_type)
```

Crime types
```{r}
#Showing distinct crimes: 22 types of crime here.
pg_county_crime_old %>% 
  distinct(crime_type)
```

Ditching records before 2019 and 2023 because we are using the Open Data portal records for 2019-2022 and using 2023 from the PG Police.

Since we're not using 2023 records, we no longer need to include the dataset from July 2023-Present.

```{r}
pg_county_crime_old <- pg_county_crime_old %>% 
  filter(year >= 2019, year <= 2022)
```

filtering for AUTO, STOLEN and "AUTO, STOLEN & RECOVERED"
```{r}
pg_stolen <- pg_county_crime_old %>% 
  filter(crime_type == "AUTO, STOLEN" | crime_type == "AUTO, STOLEN & RECOVERED" )
```

Exploratory data section

# Question: How many cars were stolen per year. Showing this in a quick visual

Goal of below code block is to show year, stolen_auto count, stolen_recovered_count and a grand_total
```{r}
pg_yearly_data <- pg_stolen %>%
  group_by(year) %>%
  summarize(
    stolen_autos = sum(crime_type == "AUTO, STOLEN")) %>% 
  add_row(year = 2023, stolen_autos = 6229) %>% 
  mutate(perc_change = (6229-3105)/3105)*100

```

quick visual
```{r}
ggplot() + 
  geom_bar(
    data=pg_yearly_data, 
    aes(x=year, weight=`stolen_autos`))+
  labs(title = "Car thefts in PG county 2019-2022",
       caption =  "Source: PG Open Data | Graph by Dylan Manfre")
```
Confirms a steady increase from 2019 to 2022. Again, we only did up to 2022 because we are using 2023 numbers from the PG police.

Monthly totals for PG county 2019-2022
```{r}
pg_monthly <- pg_stolen %>% 
  group_by(month, year) %>% 
  summarize(
    stolen_recovered = sum(crime_type == "AUTO, STOLEN & RECOVERED"),
    stolen_autos = sum(crime_type == "AUTO, STOLEN"),
    month_total = stolen_recovered + stolen_autos) %>% 
    mutate(month = month(month, label = TRUE))

```

