library
```{r}
library(tidyverse)
library(janitor)
library(lubridate)
#install.packages('tidygeocoder')
library(tidygeocoder)
#install.packages("curl")
library(curl)
library(ggplot2)
library(ggrepel)
```

loading data
```{r}
pg_county_crime_old <-  read.csv("data/pg_county_crime.csv")
updated_pg <- read_csv("data/updated_pg.csv")
```

cleaning
```{r}
pg_county_crime_old <- pg_county_crime_old %>% 
  clean_names() %>% 
 mutate(date = as.Date(date, "%m/%d/%Y"),
        year = year(date),
         month = month(date)) %>% 
  rename(crime_type = clearance_code_inc_type) %>% 
  filter(date <= "2023-07-01")

updated_pg <- updated_pg %>% 
   clean_names() %>% 
 mutate(date = as.Date(date, "%m/%d/%Y"),
        year = year(date),
         month = month(date)) %>% 
  rename(crime_type = clearance_code_inc_type)


total_pg_county_data <- bind_rows(pg_county_crime_old, updated_pg )


```

Crime types
```{r}
#Showing distinct crimes: 22 types of crime here.
total_pg_county_data %>% 
  distinct(crime_type)
```

Ditching records before 2019 and 2023 because we are using the Open Data portal records for 2019-2022 and using 2023 from the PG Police.

# UPDATE FROM NOV. 8. To check the month totals, I added 2023 back to the dataset. We are not making graphics out of these but we are using the numbers fro the story and relying on the police for the yearly 2023 number which was said in the email (6229).

```{r}
total_pg_county_data <- total_pg_county_data %>% 
  filter(year >= 2019)
```

filtering for AUTO, STOLEN.
```{r}
pg_stolen <- total_pg_county_data %>% 
  filter(crime_type == "AUTO, STOLEN" |crime_type == "AUTO, STOLEN & RECOVERED")
```

Exploratory data section

# Question: How many cars were stolen per year. Showing this in a quick visual

Goal of below code block is to show year, stolen_auto count, stolen_recovered_count and a grand_total
```{r}
pg_yearly_data <- pg_stolen %>%
  group_by(year) %>%
  summarize(
    stolen_recovered = sum(crime_type == "AUTO, STOLEN & RECOVERED"),
    stolen_autos = sum(crime_type == "AUTO, STOLEN")) %>% 
    mutate(grand_total = stolen_autos +stolen_recovered)
    #add_row(year = 2023, stolen_autos = 6229)

#This number below does not include the 2023 number from the police (6229).
pg_yearly_data %>% 
  summarize(stolen_autos = (sum(stolen_autos)))

#write_csv(pg_yearly_data, "data/yearly_total.csv")

```

quick visual
```{r}
ggplot() + 
  geom_bar(
    data=pg_yearly_data, 
    aes(x=year, weight=`grand_total`))+
  labs(title = "Car thefts in PG county 2019-2023",
       caption =  "Source: PG Open Data and Prince George's Police Department | Graph by Dylan Manfre",
       y = "Stolen Cars",
       x = "Year")
```
Confirms a steady increase from 2019 to 2022. Again, we only did up to 2022 because we are using 2023 numbers from the PG police.

Monthly totals for PG county 2019-2022
```{r}
pg_monthly <- pg_stolen %>% 
  group_by(month, year) %>% 
  summarize(
    stolen_recovered = sum(crime_type == "AUTO, STOLEN & RECOVERED"),
    stolen_autos = sum(crime_type == "AUTO, STOLEN")) %>%     
    mutate(month = month(month, label = TRUE)) %>%
    mutate(grand_total = stolen_autos + stolen_recovered)
    #pivot_wider(names_from = year, values_from = grand_total)


```

